---
title: "Twitter textual data: scraping and analyzing"
author: "Renato Russo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The 3 big tasks behind analyzing textual Twitter data
### 1. Scrape data
The first thing you should do is obtaining data. Fortunately, Twitter makes it 
relatively easy to scrape data, all you need to start is a Twitter Developer
Account so you can get access to Twitter API. Actually, there are a few different
types of Developer Account. For the purpose of this project, I am using my 
Academic Research account. You may want to check out [Twitter's Developer Portal](https://developer.twitter.com/)
for more information.
From this point on, I will assume that you have your Twitter API credentials at
hand ;-)

#### Packages needed/recommended
There are a few packages that help us retrieve data from Twitter. For this project,
I am using [Academic twitteR](https://www.research.ed.ac.uk/en/publications/academictwitter-an-r-package-to-access-the-twitter-academic-resea)

#### Set up credentials

#### Search what you want

### 2. Prepare data
#### Packages needed

#### Remove stop words

#### Remove punctuation

#### Tokenize words

#### Create DTM

### 3. Analyze data
#### Packages needed
#### Exploratory analysis

#### Natural language processing

##### Visualize topiccs

##### Visualize co-occurring terms


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.